---
title: "Reproducible Research - Course Assignment 1"
author: "DK"
date: "2/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read csv file:

```{r}
library(lubridate)
df <- read.csv("activity.csv")
df$date <- ymd(df$date)
head(df)
str(df)
```
  
  
Calculate average number of steps per day, output result in a histogram, calculate mean and median number of steps taken each day

```{r}
library(dplyr)
steps_day <- df %>% group_by(date) %>% summarize(stepsPerDay = sum(steps, na.rm = TRUE))
steps_day <- as.data.frame(steps_day)

hist(steps_day$stepsPerDay, col="red", xlab = "Steps by day", main = "Histogram of steps by day")
dev.copy(png,height = 480, width = 480, file= "plot1.png")
dev.off()

mean_steps <- mean(steps_day$stepsPerDay, na.rm  = TRUE)
median_steps <- median(steps_day$stepsPerDay, na.rm = TRUE)

```
Mean for daily steps is `r mean_steps`.  
Median for daily steps is `r median_steps`.  


Calculate avg number of steps in the 5 min interval, output result in a time series chart
```{r}
library(dplyr)
steps_interval <- df %>% group_by(interval) %>% summarize(avgStepsInt = mean(steps,na.rm=TRUE))
plot(steps_interval$interval,steps_interval$avgStepsInt, type = "l", xlab = "Interval", ylab = "Avg Daily Steps")
dev.copy(png,height = 480, width = 480, file= "plot2.png")
dev.off()
maxStepsInterval <- steps_interval[steps_interval$avgStepsInt == max(steps_interval$avgStepsInt),1]

```

Interval with max daily average number of steps is `r maxStepsInterval`.


Calculate and report the number of NA data in the dataset
```{r}
library(dplyr)
numberNAValues <- sum(is.na(df[,1]))
```
Number of intervals with NA data is `r numberNAValues`.  


Impute daily median values for NA
```{r}

df_upd <- df %>% mutate(steps_upd = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))

steps_day_upd <- df_upd %>% group_by(date) %>% summarize(stepsPerDay = sum(steps, na.rm = TRUE))

mean_upd <- mean(steps_day_upd$stepsPerDay)
median_upd <- median(steps_day_upd$stepsPerDay)

hist(steps_day_upd$stepsPerDay, col="red", xlab = "Steps by day", main = "Histogram of steps by day with imputed values")

dev.copy(png,height = 480, width = 480, file= "plot3.png")
dev.off()

```

Updated mean for daily steps is `r mean_upd`.  
Updated median for daily steps is `r median_upd`.  


Create a two-level factor variable for weekdays and weekends
```{r}
library(ggplot2)

df_upd1 <- df_upd %>% mutate(weekday = weekdays(date))

df_upd1 <- df_upd1 %>% mutate(DayOfWeek = ifelse(weekday %in% c("Saturday","Sunday"), "weekend","weekday"))
df_upd1$DayOfWeek <- as.factor(df_upd1$DayOfWeek)

steps_interval_upd <- df_upd1 %>% group_by(DayOfWeek,interval) %>% summarize(avgStepsInt = mean(steps,na.rm=TRUE))

ggplot(steps_interval_upd,aes(interval,avgStepsInt,col=DayOfWeek) )+geom_line() + labs(x = "Interval", y = "Avg Steps Per Interval")

dev.copy(png,height = 480, width = 480, file= "plot4.png")
dev.off()

```